%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Identifying validation results}
\label{sect:identifying}
To get an idea of the (meta)data necessary to identify a validation result,
consider the data, validation rules and validation results shown in
Table~\ref{tab:example1}. When a dataset is confronted with a validation rule,
there are three possible outcomes. A rule can be satisfied, yielding $1$, a
rule can be failed, yielding $0$, or a rule cannot be evaluated because
of missing data, yielding \na{}.
%
\begin{table}
\centering
\caption{Example data, validation rules, and results}
\begin{tabular}{rccccb{4cm}}
\hline
&\multicolumn{2}{c}{\textbf{Variables}}&&\multicolumn{2}{c}{\textbf{Rules}}\\
Nr  & Age  & hasjob     && \code{Age >= 0} & \code{IF Age < 15 THEN hasjob == `no'}\\
\cline{2-3}\cline{5-6}
1   & 36   & \code{yes} && 1        & \multicolumn{1}{c}{1}\\
2   & 53   & \code{NA}  && 1        & \multicolumn{1}{c}{\na{}}\\
3   & 11   & \code{yes} && 1        & \multicolumn{1}{c}{0}\\
\hline
\end{tabular}
\label{tab:example1}
\end{table}



In the example three records on age and work status are checked against two
rules: age must be larger than or equal to zero, and persons under 15 years old
cannot have a job. In each case the demand on  age can be checked and and each
record passes this test, yielding $1$ (\waar{}) as validation result. For the
second rule, the first record passes the check since age equals 36 and the
person has a job which is allowed by the rule. In the second case the job
status is not available (\na{}). Hence the rule cannot be checked and the
returned value is \na{} as well. Finally, in the third record there is an
11-year old with a job which is a combination that is not allowed by the rule,
yielding $0$ (\onwaar{}).


The above discussion leads to the following definition of a validation result
(see also \citet[Chapter 5]{zio2015methodology}.
%
\begin{definition}[Validation result]
A validation result is a value from the set $\{0,1,\na{}\}$.
\label{def:validationresult}
\end{definition}
Validation results are obtained as outcomes of evaluating a validation rule on
a data set. We also follow the common convention which identifies $0$ with
\onwaar{} and $1$ with \waar{}.  The numeric representation will make it easier
to define aggregators in later chapters. 


A recurring point of discussion is whether \na{} should be an allowed
validation result. The main other options are to either interrupt execution, or
to interpret the result as failed when one of the data points
necessary for evaluating a rule is missing. Because statistical data often
suffers from missing data the first option would yield many interruptions of a
statistical process flow, unless each rule guards against it by building in
clauses that detect missing values explicitly. The second option (\na{} implies
\onwaar{}) would yield a loss of information. More importantly, it assumes
(possibly unwarranted) that the user of the result follows this interpretation.

We deem both alternatives undesirable and therefore follow the definition
above, which also agrees with the definition of a formal data validation
function in the Methodology on Validation handbook published by the Validat
Foundation ESSnet \citep{zio2015methodology}.



With a \emph{validation event} we mean the physical activity where a dataset is
confronted with a single rule, yielding precisely one validation result. A
validation event is characterized by its inputs (the rule and the dataset it
pertains to) as well as its own metadata, including the time of execution and
the actor performing the execution. We adopt the following convention for
pairing the result of such an event with its identifying metadata.
%
\begin{definition}[confrontation] 
A \emph{confrontation} is a tuple $(e,r,d,v)$, where $e$ identifies the physical
validation event, $r$ identifies the evaluated validation rule, $d$
identifies the data points to which the result pertains, and $v$ is the
generated validation result.
\label{def:confrontation}
\end{definition}
%
We leave open the possibility that the components $e$, $r$ and $d$ consist of
informative tuples to identify subcomponents of the event, data, or rule. Note
that the above `definition' is not a definition in the mathematical sense.
Rather it is a convention that can be tested in a (software) environment where
data, rules, and validation events have been labeled and stored.

Finally we point out a subtlelty concerning the identification of data points
to which the result \emph{pertains}. In many cases, the data under scrutiny is
also the data involved in evaluating the rule. Consider a set (column) of
numerical values $\la{x}=(x_1,x_2,\ldots,x_m)$. In a structural business survey
for example, $\la{x}$ may consist of profit values for different enterprises.
The rule 
\begin{align*}
\textsf{mean}(x) \geq 0
\end{align*}
is a check on the whole column. The column of data is considered invalid if the mean
profit is negative. Here, the data used to compute the result is also the data that
is being validated. Now consider the following rule, based on a method of 
\citet{hiridoglou1986statistical}, evaluated for each value $x_j$ ($j=1,2,\ldots, m)$:
\begin{align*}
\max\left(\frac{x_j}{x^*},\frac{x^*}{x_j}\right) \leq h.
\end{align*}
Here, $x^*$ is a reference value, usually $\textsf{median}(x)$, and $h$ is a
fixed parameter. To evaluate this rule $x^*$ must be computed which implies
that the whole of $\la{x}$ must be known. This means that in a formal sense the
rule is a validation of $\la{x}$ as a whole. After all, changing any value
$x_{i}$ may influence the result even when $i\not=j$. The aim is of course to
evaluate the rule $m$ times, once for each value $x_j$ where the goal is to
evaluate the value of $x_j$, not the combination of values $\la{x}$ $m$ times.

Rather than identifying the values that are used in evaluating a rule it is
therefore stressed that the report identifies both the values used to compute
the results as well as the values that are (in)validated by the result.





