\section{Introduction}
\label{sect:introduction}

Data validation is at the core of every production chain in official
statistics.  Whether it is the input received from a survey, a bunch of data
records transferred from an administrative source or the result from some
internet scraping, data must be checked against our expectations in order to
process it and turn it into reliable statistics.  For a more detailed
explanation of data validation in general and the principles we identified in
validation, we refer to the handbook of data validation from the ESSnet
Foundation, by \citet{zio2015methodology} and the validation principles written
down by the ESS validation task force, see \citet[Chapter 4]{ess2017}.

Standardisation is important in official statistics and this also applies to
validation processes.  Especially in the case of cross-organisation validation,
where both a data producer and a data receiver check the same data against some
commonly agreed validation rules, standardisation is crucial to prevent
different interpretations of the results.  The more harmonised validation
reports are, the better the understanding across organisations is.  This has
been recognized by the ESSnet on Validation project (Validat Integration).  A
Work Package (WP2) was defined to attack this issue.  This report is one of the
deliverables of Work Package 2.  It contains the result of the work carried out
by various project partners and focusses on the standardisation of the output
of a validation process: the \emph{validation report}.

Obviously we are interested in the question what information can and should be
included in a validation report and the optimal way to express this
information.  In this document we develop a \emph{generic} structure for
expressing validation results.  We have the ambition to develop a validation
report structure that can be used in \emph{any validation task} in \emph{any
organisation} in any \emph{statistical domain}, for validation of
\emph{microdata} as well as \emph{aggregated data}.  To make it applicable in
machine to machine communication contexts as well as in human contexts we
design a machine-readable as well as a human readable format.  This report
focusses on the machine-readable version.

The approach we have taken is a combination of a top-down approach and a
bottom-up approach.  In the bottom-up approach a number of example validation
reports from member states and Eurostat were collected and studied to identify
common elements.  This resulted in a long-list of validation report elements
categorized into several classes such as rule metadata, process metadata,
aggregates etc.  This led to some thinking about the basic concepts that are
used in validation reports across the ESS.  We used the results from the
bottom-up approach to develop a more formal top-down approach expressed in this
report.  Step by step we built a validation report structure that is generic
enough to be used in a wide range of validation contexts in many institutes,
expressive enough to support the most recognized validation report elements
recognized from the bottom-up approach and flexible enough to be adopted in a
regional validation context while containing the standardized elements from the
generic format.  Validation tools or other statistical tools producing
validation reports as a side product should be able to use this structure for
their validation output.

In this deliverable we address the following aspects: In
Chapter~\ref{sect:demands} we ask ourselves what we should expect from a
generic validation report.  After some elaboration on the variety in richness
of a validation report, we define a number of generic demands. In
Chapter~\ref{sect:evaluations} the idea of validation events and 
aggregation events will be explained.
Chapter~\ref{sect:identifying} explains how to identify validation results
and other core elements to be used in the
definition of the validation report structure.
In Chapter~\ref{sect:aggregation} we dive deeper into the modeling of aggregates
to be contained in a validation report.
In Chapter~\ref{sect:basicreports} we formally define the generic validation report 
structure that can carry both basic and aggregated results.
Chapter~\ref{sect:standards} describes the technical implementation of the
concepts developed in the earlier Chapters.  
%Finally, Chapter~\ref{sect:examples} is devoted to some concrete examples.


